### 2022-03-16 paper/localpsf.tex
restructuring paper

\subsection{Existing work}

The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}. Conventionally, either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products. These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation. The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}. Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.

%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}. We build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}. A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation. In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}. But even with adaptive refinement, many matrix vector products with $\Aop$ may be required. In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\convkernel_i$ are computed with each matrix vector product.

@@

In this section we present three key concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use impulse response moments to estimate the supports of the functions $\phi_x$. Second, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation. Third, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}). The goal of our method is to compute H-matrix approximations of these matrices. 
%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.

@@

>>>>

\subsection{Background}

The most popular matrix-free operator approximation methods are based on low rank approximation \cite{CCGOPAPERS}. Conventionally, either the Lanczos method, or the randomized singular value decomposition \cite{HMTRANDOM}, or CUR approximation/skeletonization \cite{CUR,MAHONEY} are used to perform the low rank approximation using only matrix-vector products. However, methods based on low rank approximation are not suitable for high rank operators, and many operators of practical interest are high rank. In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form a low rank approximation of the Hessian. The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.

%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
Our work builds upon a history of ``product-convolution`` approximations---weighted sums of convolution operators with spatially varying weights. A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators. Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
For background, we recommend reading the following papers: \cite{PRODCONVGOOD}. We specifically build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}. A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation. In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}. But even with adaptive refinement, many matrix vector products with $\Aop$ are required. In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_i$ are computed with each matrix vector product.

@@

In this section we present three key concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use these moments extensively. Second, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation. Third, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}). The goal of our method is to compute H-matrix approximations of these matrices. 
%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.

@@

Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.
